---
title: "Final project for SA23204159"
author: "Zhijie Gong"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Final project for SA23204159}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Overview

__SA23204159__ is a simple R package developed to complete the final project for the 'Statistical Computing' course. This package include one data set, namely, _lung_ and five functions,namely _XGB_md_
(train a predictive model use XGBoost algorithm), _plot_roc_(a function used to plot roc curve), _plot_confusion_matrix_(a function used to visualize confusion matrix), _studentized_bootstrap_ci_(construct studentized bootstrap CI), _studentized_bootstrap_ci_C_(Rcpp version of studentized_bootstrap_ci)

 _lung_ is a dataset which contained information of patients who have lung cancer .This dataset is used to illustrate the how to construct a XGBoost prognostic model.The last two column is the endpoint status and survival months of patient,Status 1 represents the patient is censored at endpoint and status 2 represents the patient is dead at endpoint.The rest of colmuns are the covariate.
 
The R package 'microbenchmark' can be used to benchmark the above R and C++ functions.

Show details of dataset _lung_.

lung数据集包含了肺癌患者的部分信息，倒数第二列为患者的存活时间，单位为月，倒数第一列为患者在endpoint的状态，1代表患者在endpoint删失，2代表患者在endpoint死亡。其余列为患者的临床特征。
```{r}
library(SA23204159)
data(lung)
head(lung,5)
```

an example of _XGB_md_, _plot_roc_, _plot_confusion_matrix_
XGB_md函数用于构造肺癌患者的生存预测模型，其过程用到XGBoost算法，这个函数接受两个参数，第一个参数为n_month，代表需要预测在第几月时患者的生存状态，第二个参数为data，即构建模型所使用的数据集，这个数据集可以是lung，也可以是其他的数据集，但是要求data的倒数第一列为患者在endpoint的生存状态，倒数第二列为患者的存活时间。

plot_roc用于绘制上述模型预测结果的roc曲线。

plot_confusion_matrix用于可视化上述模型预测结果的混淆矩阵。

```{r}
library(SA23204159)
data(lung)
data=lung
##由于数据较多，从数据中随机抽取2000个样本
ind<-sample(c(1:dim(data)[1]),2000)
data<-data[ind,]
res<-XGB_md(18,data)
plot_roc(res$prob_train)
plot_confusion_matrix(res$confusion_matrix)
```
an example of _studentized_bootstrap_ci_ and _studentized_bootstrap_ci_C_,use dataset _lung_ to construct 95% CI of survival month.

studentized_bootstrap_ci函数用于构造某一变量的studentized bootstrap 95%置信区间，在本例中，使用studentized_bootstrap_ci函数构造肺癌患者存活时间的95%置信区间，

studentized_bootstrap_ci_C函数是上述函数的CPP版。

```{r}
library(SA23204159)
data(lung)
data=lung
##由于数据较多，从数据中随机抽取1000个样本
ind<-sample(c(1:dim(data)[1]),1000)
data<-data[ind,]
ci1<-studentized_bootstrap_ci(data$存活时间)
ci2<-studentized_bootstrap_ci_C(data$存活时间) ##Rcpp version
print(ci1)
print(ci2)
```


## Benchmarking _studentized_bootstrap_ci_ and _studentized_bootstrap_ci_C_

The R code for benchmarking _studentized_bootstrap_ci_ and _studentized_bootstrap_ci_C_ is as follows.

比较studentized_bootstrap_ci函数和studentized_bootstrap_ci_C的运行时间。

```{r}
library(microbenchmark)
library(SA23204159)
data(lung)
data=lung
##由于数据较多，从数据中随机抽取500个样本
ind<-sample(c(1:dim(data)[1]),500)
data<-data[ind,]
tm <- microbenchmark(
  vR = studentized_bootstrap_ci(data$存活时间),
  vC = studentized_bootstrap_ci_C(data$存活时间),   ##Rcpp version
  times = 20
)
knitr::kable(summary(tm)[,c(1,3,5,6)])
```

The results again show an evident computational speed gain of C++ against R.

